---
title: "Quarto Report"
subtitle: "UNICEF Data Analysis: Youth Unemployment and Child Immunization"
author: "Rani MedhaShree"
format: html
date: 26/04/2025
embed-resources: true
code-fold: true
toc: true
toc-title: "Contents"
toc-location: left
execute: 
  error: true
  warning: false
---

# Introduction

This report presents the data analysis of two important dimensions of human development: youth unemployment and child immunization.  
Using UNICEF-sourced data-sets, we first investigate employment patterns among youth across countries, and subsequently transition to examining trends in childhood vaccination as a proxy for healthcare access.

Both employment opportunities and healthcare delivery are crucial indicators of social development, and understanding these aspects helps build a more holistic view of progress across regions.


```{r}
#| warning: false
#| message: false
library(tidyverse)
library(plotly)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)
indicator1 <- read_csv("unicef_indicator_1.csv")
# head(indicator1)
```

## BAR CHART of Youth Unemployment by Gender
```{r}
ggplot(indicator1, aes(x = sex, y = obs_value, fill = sex)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = round(obs_value, 2)), 
            vjust = -0.5, size = 3) +
  scale_fill_manual(values=c("Male" = "royalblue", "Female" = "lightpink"))+
  theme_minimal() +
  labs(
    title = "Youth Unemployment Rate: Male vs Female",
    x = "Gender",
    y = "Unemployment Rate (%)",
    fill = "Gender"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 11)
  )
```
## Scatter Plot with Linear Regression Line
```{r}
ggplot(indicator1, aes(x = obs_value, y = time_period)) +
  geom_point(color = "red") +
  geom_smooth(method = "lm", se = FALSE, color = "black") +
  theme_minimal() +
  labs(
    title = "Observation Value vs Time Period",
    x = "Observation Value",
    y = "Year"
  )
```

## Youth Unemployment Distribution over the world
```{r}
fig <- plot_ly(indicator1,
               type = "choropleth",
               locations = ~alpha_3_code,
               locationmode = "ISO-3",
               z = ~obs_value,
               text = ~paste(
                 "<b>Country:</b>", country, "<br>",
                 "<b>Value:</b>", round(obs_value, 2)
               ),
               colorscale = "Viridis",
               marker = list(line = list(color = "white", width = 1)))
fig <- fig %>% layout(
  title = list(
    text = "<b>Youth Unemployment Rate by Country</b>",
    font = list(size = 20),
    x = 0.5
  ),
  geo = list(
    projection = list(type = "natural earth"),
    showframe = FALSE,
    showcoastlines = TRUE,
    coastlinecolor = "gray"
  ),
  margin = list(t = 80)
)
fig
```

# Human Development
After examining youth unemployment patterns, here is a bit extended analysis towards another key dimension of human development: child immunization.  
Childhood vaccination rates serve as an important proxy for healthcare access, infrastructure quality, and social well-being in countries.  
The following section visualizes trends in child immunization rates across years based on another UNICEF dataset.
```{r}
indicator2 <- read_csv("unicef_indicator_2.csv")
```
## Child Vaccination Trends
```{r}
vaccination_trend <- indicator2 %>%
  group_by(time_period) %>%
  summarise(avg_vaccination = mean(obs_value, na.rm = TRUE))
ggplot(vaccination_trend, aes(x = time_period, y = avg_vaccination)) +
  geom_line(color = "royalblue", linewidth = 1.5) +
  geom_point(color = "firebrick", size = 2) +
  theme_minimal() +
  labs(
    title = "Global Average Child Vaccination Rate Over Time",
    x = "Year",
    y = "Average Vaccination Rate (%)"
  ) +
  theme(plot.title = element_text(hjust = 0.5))

```

# Conclusion
This report presents a dual exploration of human development through employment and healthcare lenses.  
In the first story, youth unemployment patterns were analyzed, revealing notable gender disparities and regional differences based on UNICEF data.  
However, limitations such as lack of multi-year country-level data restricted deeper time trend analyses.

In the second story, child immunization rates were studied as a complementary indicator of development.  
The time series visualization highlights improvements in healthcare access globally, although regional disparities likely remain.

Together, these analyses demonstrate the interconnection of employment opportunities and healthcare outcomes in shaping broader human development trajectories.